---
title: "The Art of DAR"
description: "An (exhaustive) overview of the Days in AR metric from a healthcare reimbursement perspective."
format:
  html:
    reference-location: block
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| message: false
#| warning: false
#| echo: false
#| cache: false

knitr::opts_chunk$set(
  collapse  = FALSE,
  echo      = TRUE, 
  message   = FALSE, 
  warning   = FALSE,
  error     = TRUE,
  comment   = "",
  dpi       = 300, 
  dev       = "ragg_png",
  out.width = "100%",
  fig.align = "center",
  fig.width = 8
)
options(scipen = 999)
library(tidyverse)
library(gt)
library(forager)
```


**Days in Accounts Receivable** (also known as **DAR** or **Days in AR**) is a common financial metric belonging to a group of ratios called *efficiency ratios*. It measures the average amount of time it takes for a business to collect money owed from the responsible party for services rendered and billed.

It’s importance within the healthcare RCM sphere cannot be understated, as it acts as a sort of pressure monitor connected to the vital organs of financial health within an organization. This post is intended to be a thorough overview of everything you never knew you didn’t want to know about Days in AR. Let’s start off with some background and terminology.

## Accounts Receivable

Accounts Receivable (AR) represents money owed to the healthcare practice by patients and/or insurance carriers. The Accounts Receivable cycle begins with the delivery of service and continues until payment for the service is reconciled to a zero balance. The AR Balance is the total Gross Charges entered and/or submitted to an insurance payer or patient but not yet collected. The AR balance can only be reduced by receiving payments or by entering contractual or write-off adjustments. Typical performance statistics maintained by the AR department include Days in AR and Aging of Accounts.

## Aging of Accounts

An account is a billable episode of care. It begins to age once it is billed to an insurance company or patient. These outstanding accounts are monitored by age in 30-day increments (0 - 30, 31 - 60, 61 - 90, and so forth.) AR departments monitor the number of accounts outstanding, the dollar amount in each 30-day increment (sometimes called “bins” or “buckets”), and the responsible parties. As well, aging should be broken down by many metrics, such as Provider, Patient, Insurance Types (Commercial, Primary, Secondary, Worker’s Compensation, Managed Care), Facility, Diagnosis/Procedure code, Specialty, etc. The older the account or the longer the account remains unpaid, the less likely it will be reimbursed.

Most claims are originally billed to insurance and, until the insurance makes a payment, the responsibility for the payment continues to be with the insurance payer. After the payer makes or denies a payment (with no just cause for an appeal), the responsibility for the balance of goes to the patient (to be sent an invoice) or the physician (to be written off.)

## Days in Accounts Receivable

Days in Accounts Receivable is a mathematical formula that calculates the average number of days it takes a business to collect money that it is owed. The DAR number should be low. A high DAR number tells you that there is a problem in your revenue cycle.

::: {.callout-tip}

## The Healthcare Economic Model

To understand the healthcare version of Days in AR, you need to first understand a little bit about the environment in which it lives. Typically, companies sell their products or services directly to consumers and are paid immediately at the point of service. This is called a business-to-consumer revenue model. Another revenue model type is the business-to-business model, concerning businesses that transact primarily with other businesses. As purchases between two businesses tend to be much larger, business-to-business companies invoice on a monthly basis, meaning it will not be paid for 30-plus days.

Somewhere in between these two models is the American healthcare system. This gray area is caused by a middleman called a third party payer. Better known as an insurance company, providers enter into a contractual agreement with this third party to be paid for their services at a reduced rate. In exchange, a provider gains access to the payer’s large subscriber base. The difference between the charge a provider submits to a payer and the contracted rate he or she has agreed to is called the Contractual Adjustment, which is written off once a claim has been adjudicated.

The contractual adjustment introduces a critical inefficiency to the healthcare revenue model. Rather than prices being determined by supply and demand, they are set by individually negotiated arrangements, the end result of which is no real demand or pricing curve. This issue is much too complex for me to properly do justice to here. It is, however, an essential concept to grasp when trying to understand healthcare reimbursement. In other words, healthcare reimbursement is crazy, not you.

::: 

{{< pagebreak >}}

### Example: Your Days in AR is 95

So now you know that a low Days in AR number is good and a high number is bad. How low is good though? How high is bad? Is a DAR of zero the best? Let’s say that a practice has a Days in AR of 95 days. This means that, on average, it takes the practice 95 days to collect it’s outstanding balances. That doesn’t sound good, according to what I wrote a couple of paragraphs above. Or could that normal for that particular practice?

Let’s start with what we can say for sure about a Days in AR of 95:

   1. 95 days is roughly equal to a little over three months.
   2. Therefore, there is three months of unpaid work sitting in Accounts Receivable.
   3. As such, this practice will need three months of cash reserves on hand to meet its financial obligations.

The “ideal” Days in AR depends entirely upon the medical practice, its customers (patients), its customers’ payers, the services it supplies, the providers performing those services, and the financial situation of the practice.

## DAR Formulas

You need three numbers to calculate Days in AR:

   1. **Number of Days in the Period:** Literally the number of days in the period for which you are calculating Days in AR. DAR is typically measured in monthly and quarterly increments, so this number will usually be around 30 or 90.
   2. **Total Gross Charges:** The total dollar amount (derived from the practice’s fee schedule^[A list of fees physicians establish as the fair price for the services they provide. Keep in mind that this is not the same as a payment schedule.]) charged by the practice during the period that you are measuring.
   3. **Ending AR Balance:** The Accounts Receivable balance at the close of business on the final day of the period that you are measuring.

And the formula for Days in AR:

$$ \text{Days in AR} = \frac{\text{Ending AR Balance}}{\text{Total Gross Charges}\div \text{Number of Days in the Period}} $$

### Step-by-Step Example

Let’s say we want to calculate the Days in AR for January 2024. Our Gross Charges total is \$131,440.30 and our Ending AR Balance is \$203,460.00:

```{r}
#| label: ex_step_1
#| echo: true
ex <- dplyr::tibble(
  `End Date` = as.Date("2024-01-31"),
  `Gross Charges` = 131440.30,
  `Ending AR Balance` = 203460.00,
  `Days in Period` = lubridate::days_in_month(`End Date`)
  )

ex |> 
  gt::gt() |> 
  gt::fmt_currency(columns = 2:3, decimals = 2, incl_space = TRUE) |> 
  gt::opt_table_font(font = gt::google_font("Roboto Condensed")) |> 
  gt::opt_stylize() |> 
  gt::tab_options(
    table.width = "100%",
    quarto.disable_processing = TRUE, 
    column_labels.font.weight = "bold")
```


We have the three numbers we need, let’s plug them into our formula:

```{r}
#| label: ex_step_2
#| echo: true
ex <- ex |> 
  dplyr::mutate(
    `Avg. Daily Charge` = `Gross Charges` / `Days in Period`,
    `Avg. Days in AR` = `Ending AR Balance` / `Avg. Daily Charge`
    )

ex |> 
  gt::gt() |> 
  gt::fmt_currency(columns = c(2:3, 5), decimals = 2, incl_space = TRUE) |> 
  gt::fmt_number(columns = 6, decimals = 2) |>
  gt::opt_table_font(font = gt::google_font("Roboto Condensed")) |> 
  gt::opt_stylize() |> 
  gt::tab_options(
    table.width = "100%",
    quarto.disable_processing = TRUE, 
    column_labels.font.weight = "bold")
```


And there we have it, our Average Daily Charge is \$4,240.01 and our Days in AR is 47.99.

## Building on the DAR Formula

Now that we have a basic understanding of the formula for Days in AR, we can expand upon and rearrange it to do many other things.

One thing we’ll need to do is to declare a Days in AR target or DARt, which is the benchmark or threshold that we’d like to keep our Days in AR under for the time period being measured. For the time being, we’ll make it 35 days.

As well, to simply things a bit, I’m going to use acronyms for these long variable names. They are as follows:

```{r}
#| label: acronyms
#| echo: true
dplyr::tibble(
  `Acronym` = c("`NDiP`", "`GCt`", "`EARb`", "`ADC`", "`DAR`", "`DARt`"),
  `Variable` = c("Number of Days in Period", 
                 "Total Gross Charges", 
                 "Ending AR Balance", 
                 "Average Daily Charge", 
                 "Average Days in AR", 
                 "Average Days in AR Target")) |> 
  gt::gt(process_md = TRUE) |> 
  gt::fmt_markdown(columns = 1, md_engine = "commonmark") |>
  gt::opt_table_font(font = gt::google_font("Roboto Condensed")) |> 
  gt::opt_stylize() |> 
  gt::tab_options(
    table.width = "100%",
    quarto.disable_processing = TRUE, 
    column_labels.font.weight = "bold")
```

