---
title: "Vectors to Regex"
subtitle: "Attempting A Functional Regex Translator"
description: "Notes on building a regex translator from a vector of equal length strings to a regex pattern."
format:
  html:
    reference-location: margin
    other-links:
      - text: Regex Cheatsheet
        icon: bookmark-plus
        href: https://github.com/raredd/regex
      - text: RexEgg
        icon: bookmark-plus
        href: https://www.rexegg.com/
      - text: Regex in R
        icon: bookmark-plus
        href: https://stat.ethz.ch/R-manual/R-devel/library/base/html/regex.html
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| message: false
#| warning: false
#| echo: false
#| cache: false

knitr::opts_chunk$set(
  collapse  = FALSE,
  echo      = TRUE, 
  message   = FALSE, 
  warning   = FALSE,
  error     = TRUE,
  comment   = "",
  dpi       = 300, 
  dev       = "ragg_png",
  out.width = "100%",
  fig.align = "center",
  fig.width = 8
)
options(scipen = 999)
library(tidyverse)
library(rlang)
library(cli)

long_test <- c(
  "0214T", 
  "0215T",
  "0217T", 
  "0218T", 
  "0219T", 
  "0220T", 
  "0221T", 
  "0222T", 
  "0263T", 
  "0265T", 
  "0266T", 
  "0269T", 
  "0274T", 
  "0275T", 
  "0329T", 
  "0330T", 
  "0422T", 
  "0444T", 
  "0445T", 
  "0506T", 
  "0507T", 
  "15777", 
  "20939", 
  "22510", 
  "22511", 
  "22512", 
  "22513", 
  "22514", 
  "22515", 
  "22526", 
  "22527", 
  "27197", 
  "27198", 
  "30801", 
  "30802", 
  "31231", 
  "32673", 
  "34713", 
  "34714", 
  "34715", 
  "34716", 
  "34717", 
  "36221", 
  "34812", 
  "34820", 
  "34833", 
  "34834", 
  "35572", 
  "50300", 
  "50540", 
  "54420", 
  "54430", 
  "55200", 
  "55250", 
  "55300", 
  "58575", 
  "58600", 
  "58605", 
  "58700", 
  "58720", 
  "58800", 
  "58805", 
  "58900", 
  "58920", 
  "58925", 
  "58940", 
  "58943", 
  "61000", 
  "61001", 
  "61253", 
  "63035", 
  "63043", 
  "63044", 
  "63045", 
  "63046", 
  "63047", 
  "63048", 
  "64421", 
  "64480", 
  "64484", 
  "64491", 
  "64492", 
  "64494", 
  "64495", 
  "64634", 
  "64636", 
  "76514", 
  "92025", 
  "92081", 
  "92082", 
  "92083", 
  "92132", 
  "92133", 
  "92134", 
  "92145", 
  "92201", 
  "92202", 
  "92227", 
  "92228", 
  "92229", 
  "92235", 
  "92240", 
  "92242", 
  "95870", 
  "C7501", 
  "C7502", 
  "C7504", 
  "C7505", 
  "C9771", 
  "E0675", 
  "G0279", 
  "G0412", 
  "G0413", 
  "G0414", 
  "G0415", 
  "S2342")
```


# Links

   - [Regex Cheatsheet](https://github.com/raredd/regex)
   - [Learn Regex](https://github.com/ziishaned/learn-regex)
   - [Evaluating strings as code](https://josiahparry.com/posts/2024-06-13-eval-strings)
   - [Building rules from code operands](https://github.com/Permian-Global-Research/rsi)
   - [Filter logic](https://github.com/atorus-research/ardis/blob/main/R/denom.R)
   - [Regular-Expression.info](https://www.regular-expressions.info/tutorial.html)
   - [RexEgg](https://www.rexegg.com/)
   - [Regular Expressions as used in R](https://stat.ethz.ch/R-manual/R-devel/library/base/html/regex.html)
   - [Debuggex](https://www.debuggex.com/)
   

# Packages

   - [**textshape**](https://github.com/trinker/textshape)
   - [**stringr.plus**](https://github.com/johncassil/stringr.plus)
   - [**regexmagic**](https://github.com/jonocarroll/regexmagic)
   - [**makeunique**](https://github.com/selkamand/makeunique)
   - [**similiars**](https://github.com/davidsjoberg/similiars)
   - [**strex**](https://github.com/rorynolan/strex/)
   - [**stringfish**](https://github.com/traversc/stringfish)
   - [**collapse**](https://sebkrantz.github.io/collapse/)
   - [**kit**](https://github.com/2005m/kit)
   - [**cheapr**](https://github.com/NicChr/cheapr)
   - [**fastplyr**](https://github.com/NicChr/fastplyr)
   - [**stringi**](https://stringi.gagolewski.com/)
   - [**qdapRegex**](https://github.com/trinker/qdapRegex)
   - [**charcuterie**](https://github.com/jonocarroll/charcuterie)

# Examples

   - [mark::globbing](https://github.com/jmbarbone/mark/blob/main/R/glob.R)
   - [RUtil::make_unique_comb](https://github.com/raivokolde/RUtil/blob/master/Source/R/make_unique_comb.r)
   - [mirmisc::longest_common_substring](https://github.com/mirvie/mirmisc/blob/main/R/utils.R)
   - [hutils::longest prefix/suffix](https://github.com/HughParsonage/hutils/blob/master/R/longest_affix.R)
   - [tidyext::combn_2_col](https://github.com/m-clark/tidyext/blob/master/R/combn_2_col.R)

# Functions

```{r}
#| label: fuimus_northstar
fuimus::construct_regex(northstar::search_descriptions()$hcpcs_code)
# "^[A-CEGHJ-MP-V0-9][0-9]{3}[AFMTU0-9]$"

fuimus::construct_regex(procedural::order()$code)
# "^[B-DF-HX0-9][A-HJ-N0-Z]{5}[A-DF-HJ-NP-T0-Z]$"
#                      ^^^                 ^^^
# 0-Z should be [0-9A-Z]
```


```{r}
#| label: pos_re
#| echo: true
#| eval: true
pull_char <- function(x) {
  stringr::str_extract_all(x, stringr::regex("[A-Z]"))
}
  
pull_numb <- function(x) {
  stringr::str_extract_all(x, stringr::regex("[0-9]"))
  }

pos_re <- function(x) {
  sorted   <- stringr::str_sort(x, numeric = TRUE)
  alphabet <- purrr::list_c(pull_char(sorted))
  numbers  <- purrr::list_c(pull_numb(sorted))
  paste0(
    fuimus::collapser(alphabet), 
    fuimus::collapser(numbers)
    )
}
```


```{r}
# t(vecs) |> as.data.frame()
#   dplyr::tibble() |> 
#   janitor::clean_names()
#   
# x <- split(vecs, vecs$V1)
# run_spl |> purrr::map_dfr(fuimus::uniq_rmna)
```


```{r}
#| label: id_runs
#| echo: true
#| eval: true
id_runs <- function(x) {
  vec               <- rlang::set_names(rep(0, length(c(LETTERS, 0:9))), c(LETTERS, 0:9))
  test              <- fuimus::splitter(x)
  vecna             <- vec[test]
  vecna             <- vecna[!is.na(vecna)]
  vec[names(vecna)] <- 1

  vec_group <- dplyr::tibble(value = names(vec),
                             key = vec,
                             idx = 1:length(vec),
                             group = dplyr::consecutive_id(key)) |>
    dplyr::mutate(group_size = dplyr::n(),
                  .by = group) |>
    dplyr::filter(key == 1, group_size >= 3) |>
    dplyr::select(value, group)

  if (vctrs::vec_is_empty(vec_group)) return(x)

  xgroups <- unname(split(vec_group, vec_group$group)) |>
    purrr::map(purrr::pluck("value")) |>
    purrr::map(paste0, collapse = "") |>
    purrr::list_c()

  replacements <- dplyr::left_join(
    dplyr::slice_min(vec_group, by = group, order_by = value) |> dplyr::rename(start = value),
    dplyr::slice_max(vec_group, by = group, order_by = value) |> dplyr::rename(end = value),
    by = dplyr::join_by(group)) |>
    glue::glue_data("{start}-{end}") |>
    as.vector()

  res <- stringi::stri_replace_all_regex(x, xgroups, replacements, vectorize_all = FALSE)

  paste0("[", res, "]")
}
```

```{r}
#| label: id_runs2
#| echo: true
#| eval: true
id_runs2 <- function(x) {
  
  test <- list(
    char = purrr::map(x, pull_char) |> purrr::pluck(1, 1),
    numb = purrr::map(x, pull_numb) |> purrr::pluck(1, 1))
  
  vec <- list(
    char = rlang::set_names(rep(0, length(LETTERS)), LETTERS),
    numb = rlang::set_names(rep(0, length(as.character(0:9))), as.character(0:9)))
  
  vna <- list(
    char = vec$char[test$char],
    numb = vec$numb[test$numb])
  
  vna <- list(
    char = vna$char[!is.na(vna$char)],
    numb = vna$numb[!is.na(vna$numb)])
  
  vec$char[names(vna$char)] <- 1
  vec$numb[names(vna$numb)] <- 1
  
  group_char <- dplyr::tibble(
    value = names(vec$char),
    key = vec$char,
    idx = 1:length(vec$char),
    group = dplyr::consecutive_id(key)) |>
    dplyr::mutate(group_size = dplyr::n(), .by = group) |>
    dplyr::filter(key == 1, group_size >= 3) |>
    dplyr::select(value, group)
  
  group_numb <- dplyr::tibble(
    value = names(vec$numb),
    key = vec$numb,
    idx = 1:length(vec$numb),
    group = dplyr::consecutive_id(key)) |>
    dplyr::mutate(group_size = dplyr::n(), .by = group) |>
    dplyr::filter(key == 1, group_size >= 3) |>
    dplyr::select(value, group)
  
  xgroups_char <- unname(split(group_char, group_char$group)) |> 
    purrr::pluck(1, "value") |> 
    glue::glue_collapse(sep = "")
  
  xgroups_numb <- unname(split(group_numb, group_numb$group)) |>
    purrr::pluck(1, "value") |> 
    glue::glue_collapse(sep = "")
  
  replace_char <- dplyr::left_join(
    dplyr::slice_min(group_char, by = group, order_by = value) |> dplyr::rename(start = value),
    dplyr::slice_max(group_char, by = group, order_by = value) |> dplyr::rename(end = value),
    by = dplyr::join_by(group)) |>
    glue::glue_data("{start}-{end}") |>
    as.vector()
  
  replace_numb <- dplyr::left_join(
    dplyr::slice_min(group_numb, by = group, order_by = value) |> dplyr::rename(start = value),
    dplyr::slice_max(group_numb, by = group, order_by = value) |> dplyr::rename(end = value),
    by = dplyr::join_by(group)) |>
    glue::glue_data("{start}-{end}") |>
    as.vector()
  
  orig <- list(
    char = paste0(test$char, collapse = ""),
    numb = paste0(test$numb, collapse = ""))
  
  res <- list(
    char = stringi::stri_replace_all_regex(orig$char, xgroups_char, replace_char, vectorize_all = FALSE),
    numb = stringi::stri_replace_all_regex(orig$numb, xgroups_numb, replace_numb, vectorize_all = FALSE))
  
  paste0("[", res$char, res$numb, "]")
}
```


```{r}
#| label: construct_regex
#| echo: true
#| eval: true
construct_regex <- function(x) {

  x <- gsub(" ", "", fuimus::uniq_rmna(x))

  vecs <- stringr::str_split_fixed(x, "", n = max(collapse::vlengths(x))) |>
    as.data.frame() |>
    purrr::map(na_if_common)

  to_brackets <- vecs |>
    purrr::map(fuimus::uniq_rmna) |>
    purrr::map(pos_re)

  to_vec <- to_brackets |>
    purrr::map(id_runs)

  nobrack <- which(stringr::str_detect(to_vec, "\\[|\\]", negate = TRUE))

  to_vec[nobrack] <- purrr::map(to_vec[nobrack], \(x) paste0("[", x, "]"))

  to_vec <- purrr::list_c(to_vec)

  if (collapse::any_duplicated(to_vec)) {

    # TODO THIS MUST COME BEFORE APPLYING QUESTION MARKS
    # probably need to vectorize this, will surely
    # have more than one unique duplicate out of order

    dupe_idx <- which(collapse::fduplicated(to_vec, all = TRUE))

    rp <- paste0(to_vec[dupe_idx][1], "{", length(dupe_idx), "}")

    to_vec[dupe_idx] <- rp

    to_vec <- collapse::funique(to_vec)

  }

  x <- paste0("^", fuimus::collapser(to_vec), "$")

  return(x)
}
```

{{< pagebreak >}}

# HCPCS

```{r}
#| label: hcpcs_tests
#| echo: true
#| eval: true
hcpcs_tests <- list(
  one = c(0:9, LETTERS),               # "^[0-9A-Z]{5}$"
  two = c(as.character(90:98)),        # "^[9][0-8][0-9]{3}$"
  three = c(as.character(921:929)),    # "^[9][2][1-9][0-9]{2}$"
  four = c(as.character(9202:9208)),   # "^[9][2][0][2-8][0-9]{1}$"
  five = c(as.character(92015:92019))) # "^[9][2][0][1][5-9]$"

hcpcs_tests
```


```{r}
#| label: build_hcpcs_regex
#| echo: true
#| eval: true
build_hcpcs_regex <- function(x) {
  
  x <- gsub(" ", "", fuimus::uniq_rmna(x))
  
  vecs <- stringr::str_split_fixed(
    x, "", n = max(collapse::vlengths(x))) |>
    as.data.frame() |>
    purrr::map(fuimus::na_if_common)
  
  to_brackets <- vecs |>
    purrr::map(fuimus::uniq_rmna) |>
    purrr::map(pos_re)
  
  if (length(to_brackets) == 1) {
    to_vec <- purrr::map(to_brackets, id_runs2)
  }
  
  if (length(to_brackets) > 1) {
    to_vec <- purrr::map(to_brackets, id_runs)
  }
  
  nobrack <- which(stringr::str_detect(to_vec, "\\[|\\]", negate = TRUE))
  
  to_vec[nobrack] <- purrr::map(to_vec[nobrack], \(x) paste0("[", x, "]"))
  
  to_vec <- purrr::list_c(to_vec)
  
  if (collapse::any_duplicated(to_vec)) {
    
    # TODO probably need to vectorize this, will surely
    # have more than one unique duplicate out of order
    
    dupe_idx <- which(collapse::fduplicated(to_vec, all = TRUE))
    
    rp <- paste0(to_vec[dupe_idx][1], "{", length(dupe_idx), "}")
    
    to_vec[dupe_idx] <- rp
    
    to_vec <- fuimus::uniq_rmna(to_vec)
    
  }
  # paste0("^", fuimus::collapser(to_vec), "$")
  fuimus::collapser(to_vec)
}

build_hcpcs_regex(hcpcs_tests$one)
build_hcpcs_regex(hcpcs_tests$two)
build_hcpcs_regex(hcpcs_tests$three)
build_hcpcs_regex(hcpcs_tests$four)
build_hcpcs_regex(hcpcs_tests$five)

purrr::map_chr(long_test, build_hcpcs_regex)
```


```{r}
library(seqtrie)

tree <- RadixTree$new()
tree$insert(long_test)
results <- tree$search(long_test, max_distance = 3, mode = "global", nthreads = 4)

dplyr::filter(
  results, 
  query != target) |> 
  dplyr::tibble() |> 
  # dplyr::select(query, target) |> 
  dplyr::mutate(group = dplyr::consecutive_id(query)) |>
  tidyr::pivot_longer(
    cols = c(query, target), 
    names_to = "key", 
    values_to = "val") |>
  dplyr::arrange(group, key) |> 
  dplyr::distinct(val, .keep_all = TRUE) |> 
  print(n = 300)

tree$prefix_search("9")
```

```{r}
#| label: hcpcs_regex_long
#| echo: true
#| eval: true
x <- long_test

dataframe_to_list <- \(x) as.list(unname(x))

hcpcs_regex_long <- function(x) {
  
  x <- gsub(" ", "", fuimus::uniq_rmna(x))
  
  vecs <- stringr::str_split_fixed(
    x, "", n = max(collapse::vlengths(x))) |>
    as.data.frame()
  
  one  <- dplyr::filter(vecs, V1 %in% names(which(table(vecs$V1) == 1)))
  rest <- dplyr::filter(vecs, V1 %in% names(which(table(vecs$V1) > 1)))
  
  dplyr::mutate(rest, .id = dplyr::consecutive_id(V1), .before = 1) |> 
  # fuimus::combine(name = key, columns = c("V1", "V2"), sep = "") |>
    tidyr::pivot_longer(
    cols = !.id, 
    names_to = "key", 
    values_to = "val") |> 
    dplyr::reframe(
      .id,
      .key = readr::parse_number(key),
      val) |>
    dplyr::arrange(.id) |> 
    dplyr::group_by(.id, .key) |>
    dplyr::mutate(n = dplyr::n_distinct(val)) |>
    dplyr::ungroup() |> 
    tidyr::pivot_wider(
      names_from = n, 
      values_from = val) |>
    print(n = 500)
  
  vctrs::vec_split(rest, rest[c("V1", "V2")]) |>
    dplyr::tibble() |>
    tidyr::unpack(key) |> 
    fuimus::combine(name = key, columns = c("V1", "V2"), sep = "") |>
    tidyr::unnest(val) |> 
    dplyr::select(-c(V1, V2)) |> 
    as.data.frame() |> 
    split(rest[1])
  
  # sort(table(unname(table(vecs$V1))))
  
  # list(
  #   dplyr::filter(vecs, V1 %in% names(which(table(vecs$V1) == 1))),
  #   dplyr::filter(vecs, V1 %in% names(which(table(vecs$V1) == 5))),
  #   dplyr::filter(vecs, V1 %in% names(which(table(vecs$V1) == 11))),
  #   dplyr::filter(vecs, V1 %in% names(which(table(vecs$V1) == 15))),
  #   dplyr::filter(vecs, V1 %in% names(which(table(vecs$V1) == 17))),
  #   dplyr::filter(vecs, V1 %in% names(which(table(vecs$V1) == 19))),
  #   dplyr::filter(vecs, V1 %in% names(which(table(vecs$V1) == 21)))
  # )
  
  grps <- dplyr::mutate(
    vecs, 
    id = dplyr::consecutive_id(V1, V2)) |> 
    dplyr::add_count(id) |>
    tidyr::nest(groups = -c(id, n)) |> 
    dplyr::arrange(dplyr::desc(n))
  
  grps |> 
    dplyr::filter(n == max(n)) |> 
    dplyr::pull(groups) |> 
    purrr::pluck(1) |>
    purrr::map(fuimus::uniq_rmna) |>
    purrr::map(pos_re)
  
  to_brackets <- chops[[1]] |>
    purrr::map(fuimus::uniq_rmna) |>
    purrr::map(pos_re)
  
  if (length(to_brackets) == 1) {
    to_vec <- purrr::map(to_brackets, id_runs2)
  }
  
  if (length(to_brackets) > 1) {
    to_vec <- purrr::map(to_brackets, id_runs)
  }
  
  nobrack <- which(stringr::str_detect(to_vec, "\\[|\\]", negate = TRUE))
  
  to_vec[nobrack] <- purrr::map(to_vec[nobrack], \(x) paste0("[", x, "]"))
  
  to_vec <- purrr::list_c(to_vec)
  
  if (collapse::any_duplicated(to_vec)) {
    
    # TODO probably need to vectorize this, will surely
    # have more than one unique duplicate out of order
    
    dupe_idx <- which(collapse::fduplicated(to_vec, all = TRUE))
    
    rp <- paste0(to_vec[dupe_idx][1], "{", length(dupe_idx), "}")
    
    to_vec[dupe_idx] <- rp
    
    to_vec <- fuimus::uniq_rmna(to_vec)
    
  }
  # paste0("^", fuimus::collapser(to_vec), "$")
  fuimus::collapser(to_vec)
}

build_hcpcs_regex(hcpcs_tests$one)
build_hcpcs_regex(hcpcs_tests$two)
build_hcpcs_regex(hcpcs_tests$three)
build_hcpcs_regex(hcpcs_tests$four)
build_hcpcs_regex(hcpcs_tests$five)

purrr::map_chr(long_test, build_hcpcs_regex)
```

<br>
<br>

# Session Information

```{r}
#| label: sessioninfo
#| echo: false
#| eval: true
sessioninfo::session_info(pkgs = "loaded", info = "packages")
```
